The problem and promise of scale in multilayer networks

### 2020-02-04 paper/manuscript.Rmd
reorg panels and text refs throughout

We generated a multilayer social network consisting 16 individual nodes and two
aspects (season and landcover). Two season elementary-layers (winter and summer)
and three landcover elementary-layers (open, forest and lichen) combined for a
total of six layers. Layers were composed of individuals represented as nodes
and social associations represented as edges, for each season and landcover
combination. Proximity-based social network layers were generated using the R
package `spatsoc` [@Robitaille_2019] in `r version$version.string`
[@R_Core_Team_2019]. *FIX VANDER WAL IN SPATSOC REF*

@@

<!-- TODO: discuss layer similarity for social threshold -->
In general, individuals in multilayer networks became more connected as social scale 
increased; however, this relationship varied by habitat type and
season. As social scale increased, connective redundancy increased rapidly when
groups were formed up to approximately 100m threshold, followed by a plateau and slight
decrease at higher thresholds, presumably due to an increase in summer forest
degree at ~500m (Figure \@ref(fig:socres)b). Regardless of social scale, layer
relevance did not change for any habitat types or seasons, which suggests
habitat-by-season layers captured most, if not all, of the potential connections
an individual would have across all networks (Figure \@ref(fig:socres)c).
Finally, somewhat predictably, graph strength for winter networks increased up
to ~100-200m thresholds, followed by a plateau, while graph strength for summer
networks did not change as the social threshold increased (Figure
\@ref(fig:socres)d).  Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20-100m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50m is
likely an ideal social scale for generating social networks in caribou and
similar species.

<!-- landcover resolution
* decreasing availability of patchy lichen = no edges in these layers, though some small proportion of relocations still exist in those layers
* decreasing connective redundancy because through resolution because lichen decreased in connectivity and landcovers became more homogeneous (?)
* decreasing winter lichen relevance because the degree + number of edges decreased
--> 
<!-- TODO: discuss things in a,b,c,d order consistently. They shouldn't be presented out of order -->
Increasing the scale of landcover resolution resulted in decreased proportion of rare and
patchy habitats and low connectivity of individuals in these habitats. In
winter, layer relevance was stable for open and forest habitats but decreased
for forage habitats (Figure \@ref(fig:lcres)c). Connective redundancy decreased
with increasing landcover resolution due the decrease in connectivity of
individuals in winter forage habitats, resulting in a multilayer network where
neighbours were less frequently repeated in multiple layers <!-- TODO: I'm having a hard 
time understanding this previous sentence --> (Figure\@ref(fig:lcres)b). Layer similarity 
was generally stable for forest and open
habitats except for a contrasting increase in similarity for forest and decrease
in similarity for open habitats at 250m landcover resolution (Figure
\@ref(fig:lcres)a). Layer similarity was not possible to calculate between
winter and summer forage because, at all landcover resolutions greater than 10m
there were no individuals connected in the network layers. <!-- individuals
still used forage but in much lower proportion + lowered opportunity to
associate --> These results indicate the importance of matching landcover
resolution to scale of selection as well as ensuring the landcover product used
has a resolution fine enough to detect relevant landscape features and habitats. <!-- TODO: how? This isn't clear enough for me to follow yet -->

@@

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics generally appeared to stabilize.
Like coarse landcover resolution (above), low numbers of observations were not
sufficient for capturing individuals in all contexts and layer similarity could
not be calculated below approximately 75 observations for forest, approximately 210 
observations for openand approximately 300 observations for forage (Figure 
\@ref(fig:nobs)a). Once sufficient observations were obtained, layer similarity stabilized 
for all three landcover classes at approximately 200 observations for forest and approximately 300 observations for open and forage landcover types. Connective redundancy increased with increasing number of
observations as individuals became more connected to neighbours in multiple
layers (Figure \@ref(fig:nobs)b). Given the overall trend of these multilayer
metrics stabilizing after a certain number of observations are included, the
results suggest this sensitivity method could be useful for determining the
number of observations necessary for sufficiently describing the social dynamics
across contexts. <!-- TODO: I'm not sure this is relevant since this is the main way it is already done, can we make this thought more relevant to picking scales or multilayer networks in general? I think it's important for building robust networks -->

@@

Contrasting winter and summer seasons became more similar with increasing window
length. Layer similarity between seasons increased consistently across landcover
class (Figure \@ref(fig:winlen)a). Individual graph strength and layer relevance
increased for all summer layers as individuals became more connected with
increasing window lengths (Figure \@ref(fig:winlen)c, Figure
\@ref(fig:winlen)d). These results suggest that an increasing window length,
especially for the summer season, may be resulting in less precisely defined
seasons that are not clearly capturing periods of aggregation and
dissaggregation according to resource availability and presence of conspecifics.
Similarly,<!-- TODO: ?? -->
<!-- TODO: end with something highlighting this shows it's important to properly define your season -->

@@

(ref:socres) Varying scale in social threshold. For each social threshold (5m, 50m, 100m, 150m, 200m, 250m, 300m, 350m, 400m, 450m, 500m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an overall increase in similarity for open layers and decrease in similarity for lichen layers. <!--TODO: what about forest.--> B) Mean connective redundancy across individuals increased rapidly between 5m and 50m reaching a plateau  of about 0.65 at approximately a 50m <!--TODO:**??**--> social threshold before decreasing between 400m and 500m. C) Mean layer relevance was relatively stable across social thresholds for each season and landcover layer. D) Mean graph strength across individuals showed an increase for winter network layers at 50m through 500m while summer network layers had consistently weak connections.

@@

(ref:lcres) Variation in multilayer network metrics as a function of changes in landcover resolution, a metric which we interpret as spatial scale. For each landcover resolution (10m, 100m, 250m, 500m and 1000m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated and we examined how network metrics changed as a function of changes in spatial scale. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250m and decrease in similarity for open layers at 250m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10m landcover resolution. B) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution. C) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. D) Mean graph strength of individuals was weaker during the summer compared to the winter months.

@@

<!-- TODO: RO first sentence... -->
(ref:nobs) Varying scale in number of observations. For each iteration, a number of observations (10-485 observations in steps of 25) were selected from the complete data set to construct multilayer networks with layers defined by each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed a stabilizing layer similarity at 200 observations for forest and 300 observations for lichen and open. <!--TODO:*There were observations of individuals in lichen and open areas at lower numbers of observations during the winter, so layer similarity could not be calculated at these values.* **Brief explanation of why two landcover don't have values at lower nobs??**.--> B) Mean connective redundancy across individuals increased with increasing number of observations, which begins to plateau around 100 observations. C) Mean layer relevance across individuals increased for winter layers and slightly increased for summer layers. D) Mean graph strength across individuals varied for winter layers becoming relatively stable at 200 observations and remained weak throughout for summer layers.

@@

(ref:winlen) Varying scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed increasing similarity consistent across landcover layers with increasing window length. B) Mean connective redundancy across individuals varied dramatically depending on the length of the time window <!--TODO:**[JWT: previously this was describing fig. 3. I don't know if what I wrote best encapsulates what connnective redundancy does here. Length of time window seems to make connective redundancy highly variable as it changes.]**.--> C) Mean layer relevance generally increased during the summer months but was relatively stable across time window lengths for each landcover layer. D) Mean graph strength across individuals was relatively stable over increasing time window lengths, though there was more variation in strength between landcover types as window length increased. <!--TODO:**[see if you agree with my changes. B-D were based off fig. 3 and not relevant to this figure]**-->

@@

(ref:winpos) Varying scale in time window position. For each time window position (shifting the start date of the time window used to define seasons 1 to 48 days from the original), multilayer networks were constructed for each landcover class (forest, lichen and ope) across two seasons (summer and winter) and four network metrics calculated. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed fluctuations in layer similarity across landcover types with shifted time windows. <!--TODO:**[JWT: previously describing window length, not position. Check that you're ok with my changes]**--> B) Mean connective redundancy across individuals increased between window position 12 through 48 due to increased connectivity in the summer network layers. C) Mean layer relevance slightly decreased for winter layers and increased for summer layers, showing decreasing connectivity in shifted winter time windows and greater connectivity in summer layers at later window positions. D) Mean graph strength across individuals increased in summer network layers at window position 25 and increased between window position 20 and 40 for winter network layers.

>>>>

We generated a multilayer social network consisting 16 individual nodes and two
aspects (season and landcover). Two season elementary-layers (winter and summer)
and three landcover elementary-layers (open, forest and lichen) combined for a
total of six layers. Layers were composed of individuals represented as nodes
and social associations represented as edges, for each season and landcover
combination. Proximity-based social network layers were generated using the R
package `spatsoc` [@Robitaille_2019] in `r version$version.string`
[@R_Core_Team_2019]. <!-- TODO: FIX VANDER WAL IN SPATSOC REF -->

@@

<!-- TODO: discuss layer similarity for social threshold -->
In general, individuals in multilayer networks became more connected as social scale 
increased; however, this relationship varied by habitat type and
season. As social scale increased, connective redundancy increased rapidly when
groups were formed up to approximately 100m threshold, followed by a plateau and slight
decrease at higher thresholds, presumably due to an increase in summer forest
degree at ~500m (Figure \@ref(fig:socres)d). Regardless of social scale, layer
relevance did not change for any habitat types or seasons, which suggests
habitat-by-season layers captured most, if not all, of the potential connections
an individual would have across all networks (Figure \@ref(fig:socres)b).
Finally, somewhat predictably, graph strength for winter networks increased up
to ~100-200m thresholds, followed by a plateau, while graph strength for summer
networks did not change as the social threshold increased (Figure
\@ref(fig:socres)a).  Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20-100m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50m is
likely an ideal social scale for generating social networks in caribou and
similar species.

<!-- landcover resolution
* decreasing availability of patchy lichen = no edges in these layers, though some small proportion of relocations still exist in those layers
* decreasing connective redundancy because through resolution because lichen decreased in connectivity and landcovers became more homogeneous (?)
* decreasing winter lichen relevance because the degree + number of edges decreased
--> 
<!-- TODO: discuss things in a,b,c,d order consistently. They shouldn't be presented out of order -->
Increasing the scale of landcover resolution resulted in decreased proportion of rare and
patchy habitats and low connectivity of individuals in these habitats. In
winter, layer relevance was stable for open and forest habitats but decreased
for forage habitats (Figure \@ref(fig:lcres)b). Connective redundancy decreased
with increasing landcover resolution due the decrease in connectivity of
individuals in winter forage habitats, resulting in a multilayer network where
neighbours were less frequently repeated in multiple layers <!-- TODO: I'm having a hard 
time understanding this previous sentence --> (Figure\@ref(fig:lcres)d). Layer similarity 
was generally stable for forest and open
habitats except for a contrasting increase in similarity for forest and decrease
in similarity for open habitats at 250m landcover resolution (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
winter and summer forage because, at all landcover resolutions greater than 10m
there were no individuals connected in the network layers. <!-- individuals
still used forage but in much lower proportion + lowered opportunity to
associate --> These results indicate the importance of matching landcover
resolution to scale of selection as well as ensuring the landcover product used
has a resolution fine enough to detect relevant landscape features and habitats. <!-- TODO: how? This isn't clear enough for me to follow yet -->

@@

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics generally appeared to stabilize.
Like coarse landcover resolution (above), low numbers of observations were not
sufficient for capturing individuals in all contexts and layer similarity could
not be calculated below approximately 75 observations for forest, approximately 210 
observations for open and approximately 300 observations for forage (Figure 
\@ref(fig:nobs)c). Once sufficient observations were obtained, layer similarity stabilized 
for all three landcover classes at approximately 200 observations for forest and approximately 300 observations for open and forage landcover types. Connective redundancy increased with increasing number of
observations as individuals became more connected to neighbours in multiple
layers (Figure \@ref(fig:nobs)d). Given the overall trend of these multilayer
metrics stabilizing after a certain number of observations are included, the
results suggest this sensitivity method could be useful for determining the
number of observations necessary for sufficiently describing the social dynamics
across contexts. <!-- TODO: I'm not sure this is relevant since this is the main way it is already done, can we make this thought more relevant to picking scales or multilayer networks in general? I think it's important for building robust networks -->

@@

Contrasting winter and summer seasons became more similar with increasing window
length. Layer similarity between seasons increased consistently across landcover
class (Figure \@ref(fig:winlen)b). Individual graph strength and layer relevance
increased for all summer layers as individuals became more connected with
increasing window lengths (Figure \@ref(fig:winlen)a, Figure
\@ref(fig:winlen)b). These results suggest that an increasing window length,
especially for the summer season, may be resulting in less precisely defined
seasons that are not clearly capturing periods of aggregation and
dissaggregation according to resource availability and presence of conspecifics.
<!-- TODO: ?? -->
<!-- TODO: end with something highlighting this shows it's important to properly define your season -->

Similarly, shifting window position of the summer season resulted in an increase in number of connections between individuals. Layer relevance

@@

(ref:socres) Varying scale in social threshold. For each social threshold (5m, 50m, 100m, 150m, 200m, 250m, 300m, 350m, 400m, 450m, 500m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated.  A) Mean graph strength across individuals showed an increase for winter network layers at 50m through 500m while summer network layers had consistently weak connections. B) Mean layer relevance was relatively stable across social thresholds for each season and landcover layer. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an overall increase in similarity for open layers and decrease in similarity for lichen layers. <!--TODO: what about forest.--> D) Mean connective redundancy across individuals increased rapidly between 5m and 50m reaching a plateau  of about 0.65 at approximately a 50m <!--TODO:**??**--> social threshold before decreasing between 400m and 500m.

@@

(ref:lcres) Variation in multilayer network metrics as a function of changes in landcover resolution, a metric which we interpret as spatial scale. For each landcover resolution (10m, 100m, 250m, 500m and 1000m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated and we examined how network metrics changed as a function of changes in spatial scale. A) Mean graph strength of individuals was weaker during the summer compared to the winter months. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250m and decrease in similarity for open layers at 250m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10m landcover resolution. D) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution.

@@

<!-- TODO: RO first sentence... -->
(ref:nobs) Varying scale in number of observations. For each iteration, a number of observations (10-485 observations in steps of 25) were selected from the complete data set to construct multilayer networks with layers defined by each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated. A) Mean graph strength across individuals varied for winter layers becoming relatively stable at 200 observations and remained weak throughout for summer layers. B) Mean layer relevance across individuals increased for winter layers and slightly increased for summer layers. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed a stabilizing layer similarity at 200 observations for forest and 300 observations for lichen and open. <!--TODO:*There were observations of individuals in lichen and open areas at lower numbers of observations during the winter, so layer similarity could not be calculated at these values.* **Brief explanation of why two landcover don't have values at lower nobs??**.--> D) Mean connective redundancy across individuals increased with increasing number of observations, which begins to plateau around 100 observations.

@@

(ref:winlen) Varying scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter) and four network metrics calculated. A) Mean graph strength across individuals was relatively stable over increasing time window lengths, though there was more variation in strength between landcover types as window length increased. <!--TODO:**[see if you agree with my changes. B-D were based off fig. 3 and not relevant to this figure]**-->  B) Mean layer relevance generally increased during the summer months but was relatively stable across time window lengths for each landcover layer. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed increasing similarity consistent across landcover layers with increasing window length. D) Mean connective redundancy across individuals varied dramatically depending on the length of the time window <!--TODO:**[JWT: previously this was describing fig. 3. I don't know if what I wrote best encapsulates what connnective redundancy does here. Length of time window seems to make connective redundancy highly variable as it changes.]**.-->

@@

(ref:winpos) Varying scale in time window position. For each time window position (shifting the start date of the time window used to define seasons 1 to 48 days from the original), multilayer networks were constructed for each landcover class (forest, lichen and ope) across two seasons (summer and winter) and four network metrics calculated. A) Mean graph strength across individuals increased in summer network layers at window position 25 and increased between window position 20 and 40 for winter network layers. B) Mean layer relevance slightly decreased for winter layers and increased for summer layers, showing decreasing connectivity in shifted winter time windows and greater connectivity in summer layers at later window positions. C) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed fluctuations in layer similarity across landcover types with shifted time windows. <!--TODO:**[JWT: previously describing window length, not position. Check that you're ok with my changes]**--> D) Mean connective redundancy across individuals increased between window position 12 through 48 due to increased connectivity in the summer network layers.
